---
title: "Health_Science_HT"
author: "Haleigh Tomlin"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(network)
library(igraph)
```


colorBlindness: https://cran.r-project.org/web/packages/colorBlindness/vignettes/colorBlindness.htm

o 11-Management Occupations

o 13- Business and Financial Operations Occupations

o 15-Computer and Mathematical Occupations

o 17- Architecture and Engineering Occupations

o 19- Life, Physical, and Social Science Occupations

o 27-Art, Design, Entertainment, Sports, And Media Occupations

o 29-Healthcare Practitioners and Technical Occupations

o 35-Protective Service Occupations

o 43-Office and Administrative Support Occupations

o 45-Farming, Fishing, and Forestry Occupations

o 47-Construction and Extraction Occupations

o 49-Installation, Maintenance, and Repair Occupations

o 51-Production Occupations

o 53- Transportation and Material Moving Occupations

```{r}
hs_data <- read_xlsx("Health_Science.xlsx")
```

```{r}
table(hs_data$...6)
```

^^ I put these SOCs in emily's code.
```{r}
hs_stem_creds <- read.csv("hs_stem_credentials.csv")
```

```{r}
hs
```

joining back to the HS data.

```{r}
table(unique(hs$...6))
```



```{r}
hs <- hs_data %>% left_join(hs_stem_creds, by = c("Code" = "SOC"))
hs <- hs[-1]
hs <- hs %>% mutate("STEM" = `...6`) 
hs <- subset(hs, select = c(name = Certification.Name, Occupation, Code, STEM))
```

```{r}
hs_attributes1 <- hs %>% select(name = Occupation, Code, STEM)
hs_attributes2 <- hs %>% select(name = Certification.Name, Code, STEM)

hs_attributes1 <- hs_attributes1[!duplicated(hs_attributes1$name),]
hs_attributes2 <- hs_attributes2[!duplicated(hs_attributes2$name),]

hs_attributes <- rbind(hs_attributes1, hs_attributes2)
```

```{r}
dim(hs_attributes)
length(unique(hs$Occupation)) + length(unique(hs$Certification.Name))
```

```{r}
hs <- hs[!duplicated(hs),]
```

```{r}
table(hs$Occupation %in% hs_attributes$name)
table(hs$Certification.Name %in% hs_attributes$name)
```

```{r}
library(dplyr)
hs[which(! hs$Certification.Name %in% hs_attributes$name) ,]
hs[which(! hs$Occupation %in% hs_attributes$name) ,]
```

```{r}
health_sci <- graph.data.frame(hs, directed = F, vertices = hs_attributes)
V(health_sci)$type <- bipartite.mapping(health_sci)$type
health_sci.pr <- bipartite.projection(health_sci)
```

```{r, fig.height = 20, fig.width = 20}
plot(health_sci,
     vertex.shape = ifelse(V(health_sci)$type, "circle", "square"),
     vertex.color = ifelse(V(health_sci)$type == F, ifelse(V(health_sci)$STEM == "STW", cbPalette[5], cbPalette[1]), cbPalette[4]),
     vertex.size = ifelse(V(health_sci)$type, 3, 7),
     vertex.label = ifelse(V(health_sci)$type == F, V(health_sci)$name, NA),
     vertex.label.color = "black",
     vertex.label.dist = ifelse(V(health_sci)$type, -.75, 0),
     vertex.label.cex = ifelse(V(health_sci)$type, .75, 1),
     edge.color = cbPalette.blk[1],
     edge.width = 1.5)
title(main = "Health Sciences Credential and Occupation Network", sub = "data from O*Net OnLine")
legend("bottomleft", legend = c("Credential", "Occupation"), col = c(cbPalette[4], cbPalette[5]), pch = c(19,15), pt.cex = 1.5, bty = "n", title = "Node Type", xjust = 0)
```



Health Sciences graph instead of nursing

For Health sciences, cybersecurity, manu we will have stem/stw


```{r}
all_df <- read_xlsx("All_Career_Clusters.xlsx")
```

```{r}
all_df
```

Vicki: "Occupation clusters include occupations from different major occupation groups."

So, I should join this to the bigger DF and then I can make a drop down?

```{r}
unique(all_df$`Code`[all_df$`Career Cluster` == "Manufacturing"])
```


```{r}
table(d$SOC %in% all_df$Code)
table(d$SOCNAME %in% all_df$Occupation)
```

```{r}
test <- substr(all_df$Code, start = 1, stop = 7)
```

```{r}
table(d$SOC %in% test)
```

