---
title: "Health_Science_HT"
author: "Haleigh Tomlin"
date: "7/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(network)
library(igraph)
```


colorBlindness: https://cran.r-project.org/web/packages/colorBlindness/vignettes/colorBlindness.htm

o 11-Management Occupations

o 13- Business and Financial Operations Occupations

o 15-Computer and Mathematical Occupations

o 17- Architecture and Engineering Occupations

o 19- Life, Physical, and Social Science Occupations

o 27-Art, Design, Entertainment, Sports, And Media Occupations

o 29-Healthcare Practitioners and Technical Occupations

o 35-Protective Service Occupations

o 43-Office and Administrative Support Occupations

o 45-Farming, Fishing, and Forestry Occupations

o 47-Construction and Extraction Occupations

o 49-Installation, Maintenance, and Repair Occupations

o 51-Production Occupations

o 53- Transportation and Material Moving Occupations

```{r}
hs <- read_xlsx("Health_Science.xlsx")
```

```{r}
hs
```

go back into onet to get the credentials for these ^^ 
get emily's code to pull the credentials for STEM ones

Health Sciences graph instead of nursing

For Health sciences, cybersecurity, manu we will have stem/stw


```{r}
all_df <- read_xlsx("All_Career_Clusters.xlsx")
```

```{r}
all_df
```

Vicki: "Occupation clusters include occupations from different major occupation groups."

So, I should join this to the bigger DF and then I can make a drop down?

```{r}
unique(all_df$`Code`[all_df$`Career Cluster` == "Manufacturing"])
```


```{r}
table(d$SOC %in% all_df$Code)
table(d$SOCNAME %in% all_df$Occupation)
```

```{r}
test <- substr(all_df$Code, start = 1, stop = 7)
```

```{r}
table(d$SOC %in% test)
```

